\documentclass[article]{memoir}%probably there's a better one somewhere

\usepackage{perpage}
\MakePerPage{footnote}

\usepackage{fontspec}
\setmainfont{Coelacanth}

\usepackage[perpage]{footmisc}
\usepackage{dblfnote}
\DFNalwaysdouble

\usepackage{alphalph}

\newcounter{vnum}
\setcounter{vnum}{1}
\setcounter{chapter}{1}

\renewcommand*{\chapter}{
	\vspace{\baselineskip}
	\settowidth{\chapindent}{\chapnumfont 999}
	\noindent\llap{\makebox[\chapindent][l]{%
		\chapnumfont \thechapter}}%
	\addtocounter{chapter}{1}
	\setcounter{vnum}{1}
}
	
\newcommand{\vnum}{%
	\textsuperscript{\thevnum}%
	\addtocounter{vnum}{1}%
}
\newcommand{\infer}[1]{\textit{#1}}
\newcommand{\focus}[1]{{#1}}
\newcommand{\lx}[1]{\textit{#1}}
\newcommand{\lxx}[2]{\textit{#1} `#2'}
%\renewcommand{\thefootnote}{\alph{footnote}}

\renewcommand*{\book}[1]{%
	\makebox[\textwidth][c]{\centering \booktitlefont #1}%\textwidth could apparently be replaced with something better here 
	\setcounter{chapter}{1}
}

%TODO - merge nearby identical footnotes??

\begin{document}

\book{1 John}
	
\chapter{} \vnum What was from the beginning,\footnote{Or ‘from the authority’ (as in sovereignty, rulership)} what we heard, what we saw with our eyes, what we watched and our hands handled, about the Message\footnote{Or ‘System’, or ‘Logic’, or ‘Argument’} of life---\vnum and the life was revealed,\footnote{Or `made real', or `made conspicuous/remarkable'; and hereafter} and we have seen and give testimony, and we report also to you\footnote{Plural, and hereafter} the eternal\footnote{literally `era-long', derived from \lxx{aion}{era}; used in secular Greek literature as `eternal, perpetual' as well} life which was beside\footnote{Literally ‘up against’} the Father and was revealed to us---\vnum that which we have seen and heard, we report to you, so that you can have partnership with us. But indeed our partnership \infer{is} together with the Father, and together with His Son, Jesus the Anointed One. \vnum And we write these things to you so that our joy can be filled up.

\vnum And \focus{this} is the report that we have heard from Him, and \infer{that} we report\footnote{Or `proclaim', or `bring news of'} to you: that God is light, and \infer{as for} darkness inside Him, there is none at all.\footnote{Lit. `there is not none'; a negative concord construction} \vnum If we ever say that that we have partnership with Him and \infer{yet} walk about in the darkness, we are lying and are not practising\footnote{Or `doing', or `engaging in', or perhaps `bringing about'} the truth. \vnum Whenever, though, we are walking about in the light, as He is in the light, we have partnership with each other, and the blood of Jesus His Son cleans\footnote{Or `prunes', or `sifts'; and hereafter} us from all mistakenness. \vnum If we ever say that we do not have mistakenness, we cause ourselves to wander, and the truth is not in us. \vnum Whenever we acknowledge our mistakes, God is faithful and righteous\footnote{Or `just', or `dutiful'} so that He can release us from \infer{our} mistakes and clean us from all unrighteousness.\footnote{Or `unjustness', or `wrongdoing'} \vnum If we ever say that we have not made mistakes, we make\footnote{\textit{Poioumen}; the same verb as `practice' above} Him a liar, and His word is not in us.

\chapter{} \vnum My little children, I am writing these things to you so that you will not make mistake. And whenever anyone makes a mistake, we have an Advocate\footnote{A technical legal term} beside the Father, Jesus the righteous Anointed \infer{One}. \vnum And he is the appeasement in regard to our mistakes; not only in regard to ours, but to the entire world. \vnum And in this we know that we have known him, if we keep his commands. \vnum The one who says `I have known him' and does not keep his commandments is a liar, and in him the truth is not present. \vnum But whoever would keep his system\footnote{The same word as `message' above}, truly in him the love of God	 has been fulfilled\footnote{Or `accomplished', or `perfected'}. By this we know that we \focus{are} in him---\vnum{}the one who claims to remain in him is obligated to, just as he in that way walked about, in the same way walk about.

\vnum Loved ones, I am not writing to you a new\footnote{\lx{Kainén}, or 'fresh'; not \textit{neén} 'new, young'} commandment, but an old commandment, which you have had from the beginning.\footnote{The same 'beginning' as in v1} The old commandment is the message that you heard. \vnum On the other hand, it is a new commandment that I am writing to you, which is true in him and in you, because the darkness is passing away, and the true light is already shining. \vnum The one who claims to be in the light but hates his brother---he is in the darkness until now. \vnum The one who loves his brother---he remains in the light and there is no trap in him. \vnum But the one who hates his brother---he is in the darkness and walks about in the darkness; and he does not know where he is going off to, because the darkness has blinded his eyes.

\indent \vnum I am writing to you, children,\\
\indent \indent because \infer{your} mistakes have been forgiven \infer{for} you for the sake of his name.\footnote{i.e.\ for his reputation} \\
\indent \vnum I am writing to you, fathers,\\
\indent \indent because you have known the \infer{one who is} from the beginning.\\
\indent I am writing to you, young men,\\
\indent \indent because you have defeated the evil \infer{one}.\\
\indent I wrote to you, children,\footnote{A different word from that in v.12---here \lx{paidía}, there \lx{teknía}}\\
\indent \indent because you have known the Father.\\
\indent \vnum I wrote to you, fathers,\\
\indent \indent because you have known the \infer{one who is} from the beginning.\\
\indent I wrote to you, young men,\\
\indent \indent because you are strong,\\
\indent \indent \indent and the Message of God remains in you,\\
\indent \indent \indent and you have defeated the evil \infer{one}.

\vnum Don't esteem\footnote{\lx{Agapâte}; or perhaps `be content with', or perhaps `prize'} the world\footnote{or `the universe', and hereafter} or the \infer{thing}s\infer{ that are} in the world. Whenever anyone esteems the world, there is not the esteem\footnote{\lx{Agápe}} of the Father\footnote{Ambiguous as to whether this is the Father's esteem for him or his esteem for the Father} in him, \vnum because all that which \infer{is} in the world---the desire of\footnote{Possibly `desire for', and hereafter} the meat, and the desire of the eyes, and the the boastfulness\footnote{Or `posturing'} of life---is \focus{not} from the Father, but is \focus{from the world}. \vnum And the world is passing away, and the desire of it; however, the \infer{one who is} practising\footnote{Again, \lxx{poion}{doing}} the intention of God remains until the era.\footnote{i.e.\ `forever'; c.f.\ \lxx{aionios}{eternal} above}

\vnum Children,\footnote{\lx{Paidía}} it is the utmost period,\footnote{\lx{Hora}, which can be for any cyclical time period---hour, season, year, etc} and just as you have heard that \infer{an} anti-anointed is coming,\footnote{Or `is arriving', or `is on the way', or `is setting out'} even now many anti-anointeds have come to be, for which reason we know that it is the utmost period. \vnum They went out from us, but they were not from us\footnote{Or `of us', `part of us'}---since if they had been from us, they would have remained among us, but \infer{it is} so that they can be revealed, because\footnote{Or `that', or `with regard to the fact that', or `as is shown by the fact that'} they are not all from us.\footnote{Or `not all are from us', or `all are not from us'} \vnum But\footnote{Literally `and'} you have \infer{an} anointing from the Holy \infer{One}, and you all know.\footnote{A different verb from all others so far---\lxx{eido}{know as fact}, rather than \lxx{gignosko}{come to know, understand}---and following. Also, some manuscripts have `you know all things'.}. \vnum I didn't write to you because\footnote{Or `that', and following} you don't know the truth, but because you know it, and because any\footnote{Or `the whole'} lie is not of\footnote{Or `from'} the truth. \vnum Who is the liar, if not the \infer{one who} denies \infer{and says} that Jesus is not the Anointed \infer{One}? This is the anti-anointed---the \infer{one who} denies the Father and the Son. \vnum Everyone \infer{who} denies the Son also does not have the Father; the \infer{one who} accepts\footnote{i.e.\ admits, grants---\lxx{homologein}{say the same}} the Son also has the Father. \vnum \infer{As for} you, what you heard from the beginning, let it remain in you. If what you heard from the beginning remains in you, you also will remain in the Son and in the Father. \vnum And this is the promise\footnote{Or `announcement'} that he promised us---eternal life.

\vnum \infer{As for} these \infer{things}, I wrote \infer{them} to you about those \infer{who are} causing you to wander. \vnum And \infer{as for} you, the anointing that you obtained from him remains in you, and you have no need for anyone to teach you.	But as\footnote{Or `since, because'} the same anointing\footnote{Some manuscripts have `his anointing', which is one letter different} teaches you about all \infer{things}, and is true and is not a lie, and just as it taught you, you remain\footnote{Or imperative `remain'} in him.

\vnum And now, children\footnote{\lx{Teknia}}, remain in him, so that when\footnote{Literally \lxx{ean}{if}, but possibly used here in imitation of Hebrew \lx{im}; this usage occasionally appears when \lxx{hotan}{when, whenever} would likely have been more natural\label{eanwhen}} he is revealed we can have boldness\footnote{Or `frankness'} and not be dishonoured\footnote{Or `ashamed'} because of\footnote{literally `from'} him in his presence.\footnote{Possibly `at his arrival'} \vnum If you know\footnote{Eidete} that he is righteous\footnote{Or `just'}, you are coming to know\footnote{Ginoskete} that also everyone \infer{who} does righteousness\footnote{Or `justice'} is born out of him. %No paragraph break here, if you can figure out how to redefine the chapter command

\chapter{} \vnum Look at what kind of love\footnote{\lx{Agapen}} the Father has given us, so that we can be called children\footnote{\lx{Tekna}} of God, and \infer{so} we are! Because of this,\footnote{Or `...we are because of this.'; `because' may also be `through'} the world does not know\footnote{\lx{Ginoskei}} us, because it did not know\footnote{\lx{Egno}; the same verb} him. \vnum Loved\footnote{\lx{Agapetoi}} \infer{ones}, now we are God's children\footnote{\lx{Tekna}}, and \infer{it} has not yet been revealed what we will be. We know\footnote{\lx{Oidamen}} that when\footnote{\lx{Ean} again; see footnote \ref{eanwhen}} he is revealed, we will be like him; because\footnote{Or `that'} we will see him just as he is. \vnum And everyone \infer{who} has this hope in him purifies himself, just as he is pure.

\vnum Everyone \infer{who} performs mistakes also performs lawlessness, and mistakenness is lawlessness. \vnum And you know\footnote{Oidate} that he appeared in order to take away mistakes, and there is no mistakenness in him. \vnum Everyone \infer{who} remains in him does not make mistakes; everyone \infer{who} makes mistakes has not seen him, and has not come to know\footnote{Egnoken} him. \vnum Children\footnote{Teknia}, do not let anyone lead you astray: the one doing righteousness is righteous, just as he is righteous. \vnum The one who performs mistakes is from the slanderer\footnote{I.e.\ the devil}, because from the beginning the slanderer has been making mistakes. For this reason the Son of God was revealed, so that he could undo the works of the slanderer.

\end{document}